Investigating Determinism of Gaming Engines for Autonomous Vehicle Verification

### 2019-12-18 paper/1_Intro.tex

% Requirements of simulation based verification SBV
For games engines to offer a suitable testing environment they must: operate with a minimal \textit{reality gap} in the physical domain (vehicles, actors), be realistic enough \cite{koopman realistic enough} including suitable rendering for perception stack testing, be easy to setup, run and manipulate the environment and the temporal development of actors \cite{Ulbrich2015.}, include the provision for sensor analogues to a suitable accuracy (e.g. video, LIDAR, IR), allow hardware-in-the-loop development and provide a suitable real-time test-bed for cyber-security testing~\cite{Javaid2013}. 
%, e.g. minimum rate of frames per second (fps).
%
% Game engines for SBV
Game engines offer a simulation domain solution for the development and testing of CAVs that meet many of these requirements but many challenges still exist.
%
% Verification and Determinism 
For games engines to be useful as a verification tool then they must be deterministic. 
%
% What is a deterministic system
Deterministic, in this context, refers to the property of causality given a temporal development of events such that any state is completely determined by prior states.
%
% Why is determinism important for verification
Determinism is important for verification. Any software defects that are detected but not repeatable at a later time or by another party does not allow for suitable resolution. Furthermore, defects not found on the verification testing system may arise on a deployed system with a different hardware architecture resulting in untested and potentially unsafe vehicle behaviour.

@@

% \noindent Existing vehicle manufacturers use vehicle dynamic simulators to model the physics of each component in a system. Similarly, the use of simulation for traffic modelling is well established. Despite these proven simulation tools, non of them provide all of the required functionalities together to fully simulate CAVs\cite{FrameworkAndChallenges}.\\\\

% move to background---->>>?
% While traffic-level simulators can model road layouts, they do not require as much realism or detail as CAVs require. On the other side, vehicle dynamics simulators include some of the high fidelity physics functionalities needed for CAV testing, but they only consider the vehicle itself and perhaps the road surface and gradient; they do not model any of the road layout, signs, traffic/street lights, other street furniture or surrounding infrastructure. 
% Most critically, neither types of simulators model sensors such as cameras, lidars or radars\cite{FrameworkAndChallenges}.\\\\ 
% Consequently, one of the major routes CAVs developers are moving towards is using physics engines, these are often games engine development frameworks, since they tick all of the boxes needed to fully simulate CAVs \cite{FrameworkAndChallenges}.  
% Robopilot, Capri, Carla, Apollo, Airsim, Udacity etc. are all examples of driving simulator projects that use physics engines to support development, training, and validation of autonomous driving systems software\cite{ListOfSimulators}. \\\\
% Test-retest reliability, is a key factor in V\&V of software. Tests need to be repeatable in order for one to eliminate bugs. 
% In other words physics engines has to be deterministic, or the level of non-determinism of the engine should be low and clearly defined, for them to be used in the context of V\&V of CAV software.\\\\
% \TODO{Remove this paragraph?} Determinism of gaming (physics) engines has not attracted much attention by engines' developers nor researchers. This previously was not important since in gaming, determinism is not critical. 
% Performance, on the other hand, is a subject of more interest in that field, where developers want games to work on all platforms and still be fast to provide a positive user experience. 
% On the contrary, determinism becomes vital and performance not when such engines are used in autonomous vehicles' AI development and testing. \\\\

>>>>

% Requirements of simulation based verification SBV
For games engines to offer a suitable testing environment they must: operate with a minimal \textit{reality gap} in the physical domain (vehicles, actors), be realistic enough \cite{Koopman2018} including suitable rendering for perception stack testing, be easy to setup, run and manipulate the environment and the temporal development of actors \cite{Ulbrich2015}, include the provision for sensor analogues to a suitable accuracy (e.g. video, LIDAR, IR), allow hardware-in-the-loop development and provide a suitable real-time test-bed for cyber-security testing~\cite{Javaid2013}. 
%, e.g. minimum rate of frames per second (fps).
%
% Game engines for SBV
Game engines offer a simulation domain solution for the development and testing of CAVs that meet many of these requirements but many challenges still exist.
%
% Verification and Determinism 
For games engines to be useful as a verification tool then they must be deterministic. 
%
% What is a deterministic system
Deterministic, in this context, refers to the property of causality given a temporal development of events such that any state is completely determined by prior states.
%
Herein the term \textit{simulation variance} is used to refer to how simulation runs vary when repeated many times and specifically the variance of agent paths within a given scenario. We will adopt the terminology defined in~\cite{Ulbrich2015}, where \textit{scene} refers to all static objects including the road network, street furniture, environment conditions and a snapshot of any dynamic elements. Dynamic elements are the elements in a scene whose actions or behaviour may change over time, these are considered actors and may include other road vehicles, cyclists, pedestrians and traffic signals. The \textit{scenario} is then defined as a temporal development between several scenes which may be specified by specific goals and values. A \textit{situation} is defined as the subjective conditions and determinants for behaviour at a particular point in time.
% 
% Why is determinism important for verification
Determinism is important for verification. Any software defects that are detected but not repeatable at a later time or by another party does not allow for suitable resolution. Furthermore, defects not found on the verification testing system may arise on a deployed system with a different hardware architecture resulting in untested and potentially unsafe vehicle behaviour.

@@

% ======================
% This is all from "FrameworkAndChallenges" without using quotes 
% \noindent Existing vehicle manufacturers use vehicle dynamic simulators to model the physics of each component in a system. Similarly, the use of simulation for traffic modelling is well established. Despite these proven simulation tools, non of them provide all of the required functionalities together to fully simulate CAVs\cite{FrameworkAndChallenges}.\\\\

% While traffic-level simulators can model road layouts, they do not require as much realism or detail as CAVs require. On the other side, vehicle dynamics simulators include some of the high fidelity physics functionalities needed for CAV testing, but they only consider the vehicle itself and perhaps the road surface and gradient; they do not model any of the road layout, signs, traffic/street lights, other street furniture or surrounding infrastructure. 
% Most critically, neither types of simulators model sensors such as cameras, lidars or radars\cite{FrameworkAndChallenges}.\\\\ 
% Consequently, one of the major routes CAVs developers are moving towards is using physics engines, these are often games engine development frameworks, since they tick all of the boxes needed to fully simulate CAVs \cite{FrameworkAndChallenges}.  
% ======================

% Robopilot, Capri, Carla, Apollo, Airsim, Udacity etc. are all examples of driving simulator projects that use physics engines to support development, training, and validation of autonomous driving systems software\cite{ListOfSimulators}. \\\\
% Test-retest reliability, is a key factor in V\&V of software. Tests need to be repeatable in order for one to eliminate bugs. 
% In other words physics engines has to be deterministic, or the level of non-determinism of the engine should be low and clearly defined, for them to be used in the context of V\&V of CAV software.\\\\
% \TODO{Remove this paragraph?} Determinism of gaming (physics) engines has not attracted much attention by engines' developers nor researchers. This previously was not important since in gaming, determinism is not critical. 
% Performance, on the other hand, is a subject of more interest in that field, where developers want games to work on all platforms and still be fast to provide a positive user experience. 
% On the contrary, determinism becomes vital and performance not when such engines are used in autonomous vehicles' AI development and testing. \\\\
